M-M paper, morphological dependence, red sequence

### 2015-04-17 papers/MbhMsph/mm.tex
fanculo

\begin{abstract}
We have the best latest M-L and M-M diagrams (66 galaxies = biggest sample ever used, IR imaging = 
minimally affected by dust and star formation, 
state-of-the-art multicomponent galaxy decompositions = 
we fit bulges, disks, spiral arms, bars, rings, haloes, extended or unresolved nuclear sources and partially depleted cores).
Previous samples were overwhelmingly dominated by early-type galaxies, our sample includes 17 spiral galaxies, 
half of which have $M_{\rm BH} < 10^7 M_\odot$,
thus we can explore the low-mass end of the diagrams.
We provide a table with spheroid luminosities, spheroid stellar masses and galaxy luminosities. 
We give updated linear regressions for Sersic/core-Sersic galaxies and for different morphological types (E/S0, Sp). 
Sersic and core-Sersic galaxies have slopes consistent with each other (contrary to what claimed by GS13). 
In the M-M diagram, bulges of early type galaxies have a slope of $\sim$1, while bulges of spiral galaxies have a slope of 2-3.
Linear regressions that we provide are symmetrical (to be compared with theoretical expectations) 
and non-symmetrical (to be used to predict BH masses).
Contrary to a previous claim, 
the bulge luminosity appears to be a better indicator of the BH mass than the galaxy luminosity.
only for early type galaxies one can use the tot luminosity as bh mass indicator {\bf check and expand}
Bulges with $n<2$, claimed by some to be pseudobulges, are not offset to lower BH masses from the correlation defined by bulges with $n>2$.

@@

%\input{table_sample.tex}

@@

\subsection{Black hole mass -- spheroid luminosity}
We show the $M_{\rm BH} - L_{\rm sph}$ diagram in Figure \ref{fig:mbhmagsph}. \\
S\'ersic and core-S\'ersic spheroids have slopes consistent with each other (within their $1\sigma$ uncertainties), 
in disagreement with the findings of GS13. 
The slope that we obtained for core-S\'ersic spheroids ($M_{\rm BH} \propto L_{\rm sph}^{1.2 \pm 0.2}$) 
is consistent with the slope reported by GS13 in the $K_s$-band ($M_{\rm BH} \propto L_{\rm sph}^{1.1 \pm 0.2}$). 
Instead, the slope that we determined for S\'ersic spheroids ($M_{\rm BH} \propto L_{\rm sph}^{1.5 \pm 0.2}$) 
is significantly shallower than that found by GS13 ($M_{\rm BH} \propto L_{\rm sph}^{2.7 \pm 0.5}$). 
Although the S\'ersic/core-S\'ersic classification used by GS13 slightly differs\footnote{which galaxies?} from the classification used here, 
the main cause of such inconsistency is that the bulge-to-total ratios obtained from our galaxy decompositions 
are different from those assumed by GS13 to convert galaxy luminosities into bulge luminosities.
Our bulge-to-total ratios for low-luminosity S\'ersic spheroids ($MAG_{\rm sph} \gtrsim -22 \rm~mag$) 
are smaller than those used by GS13. 
The host galaxies of such bulges are late-type, spiral galaxies, 
which typically present a complex morphology (bars, double bars, embedded disks, nuclear components, etc).
Our sophisticated galaxy models account for the extra components, 
while the bulge-to-total ratios of GS13 were derived from simple bulge/disk decompositions 
which overestimated the bulge luminosity.
This results in our bulge magnitudes being on average $\sim$$1\rm~mag$ fainter.
On the other side, our bulge-to-total ratios for high-luminosity S\'ersic spheroids ($MAG_{\rm sph} \lesssim -24 \rm~mag$) 
are on average larger than those adopted by GS13.
In this case, the host systems are early-type elliptical/lenticular galaxies that feature intermediate-scale disks\footnote{explain}.
Past bulge/disk decompositions failed to correctly identify the extent of such disks and treated them as large-scale disks, 
thus underestimating the bulge luminosity.
Our bulge magnitudes for such spheroids are on average $\sim$$1\rm~mag$ brighter. \\
We have seen that the change in slope of the $M_{\rm BH} - L_{\rm sph}$ correlation -- 
which is expected for consistency with other scaling relations 
(a single power-law $M_{\rm BH} - \sigma$ correlation and a double power-law $L_{\rm sph} - \sigma$ correlation) -- 
cannot be attributed to the division between the two populations of S\'ersic and core-S\'ersic spheroids.
We now test a new hypothesis, that is to say the change in slope is to be ascribed to the different formation mechanisms of early- and late-type galaxies. 
If this hypothesis is correct, 
the spheroids of early-type (elliptical and lenticular) galaxies will follow $M_{\rm BH} \propto L_{\rm sph}^{\sim 1}$, 
whereas the spheroids of late-type (spiral) galaxies will have $M_{\rm BH} \propto L_{\rm sph}^{\sim 2.5}$.
First, we checked that the slopes of the bulges of elliptical and lenticular galaxies, separately, are consistent with each other, 
and thus, together, they define a single \emph{red sequence} in the $M_{\rm BH} - L_{\rm sph}$ diagram. 
We then fit the bulges of early- and late-type galaxies with two separate log-linear regressions, 
and obtained $M_{\rm BH} \propto L_{\rm sph}^{1 \pm 0.1}$ and $M_{\rm BH} \propto L_{\rm sph}^{2.75 \pm 0.75}$ respectively, 
in excellent agreement with the theoretical expectations of our hypothesis.
{\bf vertical scatter along the early type sequence is constant?
if 2 overmassive bhs removed, what happens?}

In Figure \ref{fig:pseudob}, we show the distribution of spheroid S\'ersic indices\footnote{The spheroid S\'ersic indices 
are taken from our one-dimensional fits of the major-axis surface brightness profiles of galaxies (\emph{Paper I}).}, 
$n_{\rm sph}$, in the $M_{\rm BH} - L_{\rm sph}$ diagram.
Our aim is to check whether bulges with $n_{\rm sph}<2$, claimed by some to be pseudobulges (e.g.~ cite), 
are offset to lower black hole masses from the correlation defined by (classical) bulges with $n_{\rm sph}>2$, 
as reported by \cite{sani2011} {\bf who else?}. 
To do this, we fit a symmetrical linear regression to the bulges that have $n_{\rm sph}>2$ 
and we compute the vertical offset of all bulges from the regression. 
In the inset of Figure \ref{fig:pseudob}, we plot the vertical offset against $n_{\rm sph}$. 
From a visual inspection, bulges with $n_{\rm sph}<2$ do not appear to be offset from the correlation traced by bulges with $n_{\rm sph}>2$.

@@

>>>>

\begin{abstract}
We present the latest scaling relations between black hole mass, $M_{\rm BH}$, 
and spheroid luminosity, $L_{\rm sph}$, spheroid stellar mass, $M_{\rm *,sph}$, and galaxy luminosity, $L_{\rm gal}$,
built upon an unprecedented high-quality data set (provided here in a ready-to-use table).
Our sample, the largest ever used, counts 66 local galaxies with a dynamical measurement of $M_{\rm BH}$.
Spheroid luminosities come from high signal-to-noise $3.6\rm~\mu m$ \emph{Spitzer} satellite imagery, 
which is minimally affected by dust and star formation. 
These luminosities were derived from our state-of-the-art multicomponent galaxy decompositions, 
that take into account bulges, disks, bars, spiral arms, rings, haloes, extended or unresolved nuclear sources and partially depleted cores, 
and that -- for the first time -- were checked to be consistent with the galaxy kinematics. 
Previous studies used galaxy samples that were overwhelmingly dominated by high-mass, early-type objects.
Instead, our sample includes 17 spiral galaxies, half of which have $M_{\rm BH} < 10^7\rm~M_\odot$, 
and allows us to investigate the poorly studied low-mass end of the correlations.
We provide updated linear regressions 
(symmetrical, to be compared with theoretical expectations, and non-symmetrical, to be used to predict black hole masses) 
for S\'ersic/core-S\'ersic galaxies and for different morphological types (E/S0, Sp). 
The bulges of early-type (E/S0) galaxies follow $M_{\rm BH} \propto M_{\rm *,sph}^{1.0 \pm 0.1}$, 
as expected from a dry-merging formation scenario,
and define a tight \emph{red sequence} with intrinsic scatter {\bf ???}.
On the other hand, the bulges of late-type (Sp) galaxies define a much steeper \emph{blue sequence}, 
with $M_{\rm BH} \propto M_{\rm *,sph}^{3.0 \pm 1.3}$, 
indicating that gas-rich processes feed the black hole three times more efficiently than the host bulge. 
Disproving some previous claims, we also conclude that: 1) S\'ersic and core-S\'ersic galaxies do not define two distinct sequences; 
2) bulges with S\'ersic index $n_{\rm sph}<2$, argued by some to be pseudo-bulges, 
are not offset to lower $M_{\rm BH}$ from the correlation defined by bulges with $n_{\rm sph}>2$; 
3) $L_{\rm sph}$ and $L_{\rm gal}$ correlate equally well with $M_{\rm BH}$ only for early-type galaxies.

@@

\input{table_sample.tex}

@@

\subsection{Black hole mass -- spheroid luminosity}
We show the $M_{\rm BH} - L_{\rm sph}$ diagram in Figure \ref{fig:mbhmagsph}. \\
S\'ersic and core-S\'ersic spheroids have slopes consistent with each other (within their $1\sigma$ uncertainties), 
in disagreement with the findings of GS13. 
The slope that we obtained for core-S\'ersic spheroids ($M_{\rm BH} \propto L_{\rm sph}^{1.2 \pm 0.2}$) 
is consistent with the slope reported by GS13 in the $K_s$-band ($M_{\rm BH} \propto L_{\rm sph}^{1.1 \pm 0.2}$). 
Instead, the slope that we determined for S\'ersic spheroids ($M_{\rm BH} \propto L_{\rm sph}^{1.5 \pm 0.2}$) 
is significantly shallower than that found by GS13 ($M_{\rm BH} \propto L_{\rm sph}^{2.7 \pm 0.5}$). 
Although the S\'ersic/core-S\'ersic classification used by GS13 slightly differs\footnote{which galaxies?} from the classification used here, 
the main cause of such inconsistency is that the bulge-to-total ratios obtained from our galaxy decompositions 
are different from those assumed by GS13 to convert galaxy luminosities into bulge luminosities.
Our bulge-to-total ratios for low-luminosity S\'ersic spheroids ($MAG_{\rm sph} \gtrsim -22 \rm~mag$) 
are smaller than those used by GS13. 
The host galaxies of such bulges are late-type, spiral galaxies, 
which typically present a complex morphology (bars, double bars, embedded disks, nuclear components, etc).
Our sophisticated galaxy models account for the extra components, 
while the bulge-to-total ratios of GS13 were derived from simple bulge/disk decompositions 
which overestimated the bulge luminosity.
This results in our bulge magnitudes being on average $\sim$$1\rm~mag$ fainter.
On the other side, our bulge-to-total ratios for high-luminosity S\'ersic spheroids ($MAG_{\rm sph} \lesssim -24 \rm~mag$) 
are on average larger than those adopted by GS13.
In this case, the host systems are early-type elliptical/lenticular galaxies that feature intermediate-scale disks\footnote{explain}.
Past bulge/disk decompositions failed to correctly identify the extent of such disks and treated them as large-scale disks, 
thus underestimating the bulge luminosity.
The magnitudes that we obtained for such spheroids are on average $\sim$$1\rm~mag$ brighter. \\
We have seen that the change in slope of the $M_{\rm BH} - L_{\rm sph}$ correlation -- 
which is expected for consistency with other scaling relations 
(a single power-law $M_{\rm BH} - \sigma$ correlation and a double power-law $L_{\rm sph} - \sigma$ correlation) -- 
cannot be attributed to the division between the two populations of S\'ersic and core-S\'ersic spheroids.
We now test a new hypothesis, that is to say the change in slope is to be ascribed to the different formation mechanisms of early- and late-type galaxies. 
If this hypothesis is correct, 
the spheroids of early-type (elliptical and lenticular) galaxies will follow $M_{\rm BH} \propto L_{\rm sph}^{\sim 1}$, 
whereas the spheroids of late-type (spiral) galaxies will have $M_{\rm BH} \propto L_{\rm sph}^{\sim 2.5}$.
First, we checked that the slopes of the bulges of elliptical and lenticular galaxies, separately, are consistent with each other, 
and thus, together, they define a single \emph{red sequence} in the $M_{\rm BH} - L_{\rm sph}$ diagram. 
We then fit the bulges of early- and late-type galaxies with two separate log-linear regressions, 
and obtained $M_{\rm BH} \propto L_{\rm sph}^{1 \pm 0.1}$ and $M_{\rm BH} \propto L_{\rm sph}^{2.75 \pm 0.75}$ respectively, 
in excellent agreement with the theoretical expectations of our hypothesis.
{\bf vertical scatter along the early type sequence is constant?
if 2 overmassive bhs removed, what happens?}

{\bf coming now to figure 1: if I look at the blue points it is difficult to believe that they define a linear relation, 
their magnitude range is too small and it could just be that at low galaxy luminosities the scatter of the MBH-L relation increases. 
This is what most models predict. Do the galaxies below the correlation have some special property? 
Could it just be that spiral bulges simple do not follow a tight relation with BH mass?

Giulia, I think you already plan to do this, but can you include the error bars in all places where a relation is presented, so that readers know if there is a relation or not.
Remember also Ewan's remark about the unsuitability of the Pearson and Spearman correlation coefficients because they do not take into account the error bars on our data - 
so don't include those simple coefficients.  Actually, it would be worth pointing out to readers that a visual inspection of the plotted data requires 
the viewer to take into account the error bars when judging-by-eye the strength of a correlation. 
Maybe then mention in the following sentence that this is why the correlation coefficients are not applicable here. }

In Figure \ref{fig:pseudob}, we show the distribution of spheroid S\'ersic indices\footnote{The spheroid S\'ersic indices 
are taken from our one-dimensional fits of the galaxy major-axis surface brightness profiles (\emph{Paper I}).}, 
$n_{\rm sph}$, in the $M_{\rm BH} - L_{\rm sph}$ diagram.
Our aim is to check whether bulges with $n_{\rm sph}<2$, claimed by some to be pseudobulges (e.g.~ cite), 
are offset to lower black hole masses from the correlation defined by (classical) bulges with $n_{\rm sph}>2$, 
as reported by \cite{sani2011} {\bf who else?}. 
To do this, we fit a symmetrical linear regression to the bulges that have $n_{\rm sph}>2$ 
and we compute the vertical offset of all bulges from the regression. 
In the inset of Figure \ref{fig:pseudob}, we plot the vertical offset against $n_{\rm sph}$. 
From a visual inspection, bulges with $n_{\rm sph}<2$ do not appear to be offset from the correlation traced by bulges with $n_{\rm sph}>2$.

{\bf As you probably know, and as explained in many of my papers (e.g. http://arxiv.org/abs/1501.02937 ; http://arxiv.org/abs/1108.0997 ; http://arxiv.org/abs/1311.7207 ), 
pseudobulges are notoriously hard to identify.  It would be a mistake to have a column labelled "pseudobulge".   
However a galaxy classification column, indicating the presence of a bar, is good to include.  
While we do not wish to publish the Sersic indices here (that will be kept for a different paper), 
we could have a column indicating if n is above or below 2 - given that some people mistakingly think that they can use this as a dividing point. 
Giulia, can you also include a nearby reference to Graham (2013, http://arxiv.org/abs/1108.0997) so that people can read and understand why $n < 2$ 
does not necessarily equate to a pseudobulge. }

@@

errors on gal mag??????????????? 0.2 constant
